---
title: "Many_Analysts"
author: 'Analyst: Sabrina McNew'
date: "7/10/2020"
output:  
  html_document:
    toc: true
    toc_float: true
    theme: cerulean
    code_folding: hide
---

```{r setup, include = FALSE, message = FALSE, fig.show = 'hold'}
  knitr::opts_chunk$set(echo = TRUE)
```

#### Project info: https://osf.io/34fzc/
#### Contact: sm983@cornell.edu
  
### Project goals
The goal of this project is to analyze a dataset of breeding blue tits 
in order to answer the question "does competition affect growth of tit nestlings?"

### Project approach outline (updated July 10, 2020)
Use linear models to test for sources of variation in tarsus length, nestling 
mass, and fledging success. Covariates include, treatment (increased, decreased,
or control brood size), brood size, lay date, nestling sex, whether the nestling 
was transported into the nest or not. Random effects include: parent ids and geographical
identifiers/structures of the nest, year. 

Secondarily I tested whether the *variance* in tarsus length and width within 
a brood varies with clutch size or treatment. 

### Methods: Analyses (as written for a manuscript) 
#### Data verification 
Data were checked for completeness and exploratory analyses were conducted following
best-practices described by Zuur et al. (2010 Methods Ecol Evol). Briefly, 
I plotted key predictor and response variables to visually inspect their distribution.
I looked to make sure they were uniformly or normally distributed (depending on 
common-sense expectation of the variable) and looked for outliers that would 
indicate mis-entry typos etc. Second, I looked for homogeneity in the variance 
of different treatment groups. For both tarsus length and mass there was slightly
more variance in brood-increased treatment compared to control and brood-reduced treatments. 
However, overall the differences were minor. Finally, I checked for normality 
in continuously distributed response variables, such as tarsus. There were slight 
departures from normality in both the control and brood-increased treatment 
groups; however, these departures are common in ecological datasets and linear 
models are fairly robust to small departures in normality (Box 1976, J. Am. Stat Assoc.). 

#### Analysis
```{r, echo = F, message=F}
library(dplyr)
library(lme4)
library(lmerTest)
library(sjPlot)
library(reshape2)
```

##### Individual nestling growth and survival
I used linear mixed models (LMMs) and generalized linear mixed models (GLMMs) to test for 
differences among treatments in nestling size (day 14 tarsus length and mass) 
and fitness (fledging success and next-year survival). Tarsus length and mass
were modeled with a Gaussian distribution (i.e. LMM); fledging success and 
next-year survival were modeled with a binomial distribution (i.e. logistic GLMM).  

In addition to treatment, models included a suite of covariates
that also could have influenced nestling size and fitness. Full models included 
the fixed effects of treatment, hatch nest lay date, whether the nestling was in
its hatch nest or whether it had been experimentally moved to another nest, the 
clutch size of the nest from which the nestling hatched, and the molecular sex 
of the nestling. Non-significant factors were removed from the final model. 
Models also included the random effects of hatch year, the rearing nest id, rearing 
nestbox "nested" within nesting area, and the id of the rearing mother, rearing father, 
and genetic mother and genetic father.

I secondarily repeated these analyses by replacing the fixed effect of treatment
with the rearing nest clutch size in models of nestling tarsus and mass. Because
the brood size manipulation was designed to increase variation in brood size in 
a controlled and unbiased way, I used the post-manipulation brood size to more directly investigate
the relationship between the number of nestlings in the nest and nestling growth. 

##### Reproductive success by brood 
Second I used linear mixed effects models to study the effects of treatment 
on brood-level fledging success and variance in nestling quality within the brood.
If competition affects nestling growth, then nests in the brood-enlarged treatment
will have more variance within the brood in nestling mass and tarsus length, and
will have lower percent fledging success.  
Mass 

As a preliminary control to test whether variance is positively correlated with 
brood size by chance, I simulated data for 130 nests ranging in clutch size 
from 4 to 16. I then simulated masses for each nestling assuming that masses were
normally distributed and then calculated the variance in mass for each clutch.
Finally, I used a linear model to test for a significant relationship between 
brood size and variance in mass for the brood. In 1000 simulations the p values were slightly right-
skewed, but overall were fairly uniformly distributed. Thus, there does not 
seem to be a null expectation that variance should increase with brood size. 

```{r, echo=FALSE}
pvals <- NULL
for (i in 1:1000) {
  ns <- seq(from = 4, to = 16) %>% rep(.,10)
  ys <- lapply(ns, function(x) rnorm(x)) %>% lapply(., function(x) var(x)) %>% unlist()
  pvals[i] <- summary(lm(ys ~ ns))$coefficients[2,4]
}
hist(pvals) #looks pretty uniform (i.e. no consistent relationship between n and var, so let's forge ahead)

```

Then, I used LMMs to test for an effect of treatment on brood mass variance and 
brood tarsus length variance. Models additionally included the fixed effect 
of rearing nest lay date as well as the random effects of year, rearing nest parent 
ids, and rearing nest box nested within rearing nest area. 

Finally, I tested for a difference in fledging success among treatments 
using a binomial GLMM where the response was the paired variable of the number of 
nestlings that fledged and the number that died for each nest. The model included 
the fixed effects of rearing nest lay date as well as the random effects of year, rearing nest parent 
ids, and rearing nest box nested within rearing nest area. 

Several mixed-effects models reported the "boundary (singular) fit: see ?isSingular"
warning. Singular fits can occur when the random effects structure is too complicated
for the data. Being naive to the system, I included the full suite of random effects 
that could create structure in the data. However, someone more familiar with the 
system could potentially improve models by removing random effects that are unlikely
to affect response variables. 

### Results

#### Individual nestling success 
```{r, echo = T, cache = T}
#Load data 
tit <- read.csv("data/blue_tit_data_updated_2020-04-18.csv")

# Turn these integer columns into factors
tit <- tit %>%  mutate_at(vars(hatch_year,
                        Extra.pair_paternity,
                        rear_nest_trt,
                        home_or_away,
                        chick_sex_molec,
                        chick_survival_to_first_breed_season

), as.factor)
tit$rear_nest_trt <- relevel(tit$rear_nest_trt, ref = "7")

```

Nestling mass at day 14 was significantly lower in the brood-increased treatment 
compared to control nestlings (estimated difference in means: -0.91g; 
95%CI -1.18 - -0.64g);  p < 0.001; Table 1). However, nestling mass in the brood-reduced
treatment was not significantly different from controls (p = 0.69; Table 1). 
Hatch date of the nest, whether the nestling was reared in its natal brood or was 
transplanted, hatch nest clutch size, and nestling sex all also had significant 
effects on nestling mass (Table 1). 

```{r, warning=FALSE, cache=T}
# Table 1. Code and results for nestling mass model
mod1 <-(
  lmer(
    day_14_weight ~
      rear_nest_trt +
      hatch_nest_LD +
      home_or_away +
      hatch_nest_CS +
      chick_sex_molec +
      (1 | hatch_year) + (1 | rear_nest_breed_ID) +
      (1 | rear_mom_Ring) + (1 | rear_dad_Ring) +
      (1 | hatch_mom_Ring) + (1 | genetic_dad_ring_.WP_or_EP.) +
      (1 | rear_area / rear_Box),
    data = tit
  )
)
tab_model(mod1)
```
***

Nestling tarsus length at day 14 was also significantly smaller in enlarged broods,
compared to controls, though again there was no difference between reduced broods 
and controls (Table 2). Hatch nest lay date and hatch nest clutch size had no
significant effect on tarsus length; however, transplanted nestlings and male 
nestlings had significantly longer tarsi than nestlings reared in their hatch nest
and female nestlings (Table 2). 

```{r, warnings = F, cache = T, echo = F}
mod2 <- (lmer(
    day_14_tarsus_length ~
      rear_nest_trt +
      #hatch_nest_LD + #not sig
      home_or_away +
      #hatch_nest_CS + #not sig
      chick_sex_molec +
      (1 | rear_mom_Ring) + (1 | rear_dad_Ring) +
      (1 | hatch_mom_Ring) + (1 | genetic_dad_ring_.WP_or_EP.) +
      (1 |hatch_year) +(1|rear_nest_breed_ID) +
      (1 | hatch_Area / hatch_Box), data = tit
  )
)
tab_model(mod2, title ="Table 2.")
```
***

Nestlings were slightly more likely to survive to their first breeding season
in the brood-reduced treatment (Odds ratio estimate: 1.70, 95%CI 1.00 - 2.91; p = 0.05,
Table 3). However, there was no difference in between enlarged broods and controls. 

```{r, echo = F, cache = T, warning = F }
mod3 <- (
  glmer(
    chick_survival_to_first_breed_season ~ 
      rear_nest_trt + 
      hatch_nest_LD +
      #home_or_away + 
      #hatch_nest_CS + 
      #chick_sex_molec +
      (1 | rear_mom_Ring) + (1 | rear_dad_Ring) +
      (1 | hatch_mom_Ring) + (1 | genetic_dad_ring_.WP_or_EP.) +
      (1 | hatch_year) + (1 | rear_nest_breed_ID) +
      (1 | hatch_Area / hatch_Box),
    data = tit,
    family = "binomial"
  )
)
tab_model(mod3, title = "Table 3.")
```
***
#### Using brood size as a predictor

Models using the number of nestlings post-manipulation revealed similar effects 
on day 14 nestling mass and tarsus size. The number of nestlings in the brood was 
negatively associated with day 14 weight (Table 4). 
```{r, cache=T, echo = F, warning= F}
mod4 <- (
  lmer(
    day_14_weight ~
      rear_Cs_at_start_of_rearing +
      hatch_nest_LD +
      home_or_away +
      #hatch_nest_CS + #not significant
      chick_sex_molec +
      (1 | hatch_year) + (1 | rear_nest_breed_ID) +
      (1 | rear_mom_Ring) + (1 | rear_dad_Ring) +
      (1 | hatch_mom_Ring) + (1 | genetic_dad_ring_.WP_or_EP.) +
      (1 | rear_area / rear_Box),
    data = tit
  )
)
tab_model(mod4, title = "Table 4.")
```
***
Likewise, the number of nestlings in the brood was negatively associated with 
day 14 tarsus length (Table 5). 
```{r, cache=T, echo = F, warning= F}
mod5 <- (lmer(
  day_14_tarsus_length ~
    rear_Cs_at_start_of_rearing +
    #hatch_nest_LD + #not sig
    home_or_away +
    #hatch_nest_CS + #not sig
    chick_sex_molec +
    (1 | rear_mom_Ring) + (1 | rear_dad_Ring) +
    (1 | hatch_mom_Ring) + (1 | genetic_dad_ring_.WP_or_EP.) +
    (1 |hatch_year) +(1|rear_nest_breed_ID) +
    (1 | hatch_Area / hatch_Box), data = tit
)
)
tab_model(mod5, title = "Table 5.")
```
***

#### Results by nest

```{r, cache = T, warning = F}
# Reshape data 
nest <- dplyr::select(tit, hatch_year, rear_nest_breed_ID, rear_area, rear_Box, rear_mom_Ring, rear_dad_Ring,
               rear_nest_trt, rear_nest_LD, rear_nest_CS,rear_d0_rear_nest_brood_size,
               rear_Cs_out, rear_Cs_in, net_rearing_manipulation, rear_Cs_at_start_of_rearing,
               d14_rear_nest_brood_size, number_chicks_fledged_from_rear_nest) %>% distinct()
nest$died <- nest$rear_Cs_at_start_of_rearing - nest$number_chicks_fledged_from_rear_nest #change in brood size from start to d 14; i.e. dead nestlings

nest[duplicated(nest$rear_nest_breed_ID),] # no duplicates = no issues in data entry

variances <- aggregate(day_14_weight ~ rear_nest_breed_ID, tit, var)
variances_tar <- aggregate(day_14_tarsus_length ~ rear_nest_breed_ID, tit, var)
variances <- merge(variances, variances_tar)
names(variances)[2:3] <- c("weight_var", "tarsus_var")
nest <- merge(nest,variances)
head(nest)
```

There was significantly more variance in nestling mass among enlarged broods 
compared to controls; however, there was no difference between reduced broods and control (Table 6).
```{r, cache = T, warning = F, echo = F}
mod6 <- (lmer(weight_var~ rear_nest_trt +
               rear_nest_LD +
               (1 | rear_mom_Ring) + (1 | rear_dad_Ring) +
               (1 | hatch_year) +
               (1 | rear_area / rear_Box),
             data = nest,))
tab_model(mod6, title = "Table 6")
```
***

Likewise, there was more variance in tarsus length among enlarged broods compared
to controls but no significant difference between reduced broods and controls (Table 7). 

```{r, cache = T, warning = F, echo = F}
mod7 <- (lmer(tarsus_var~ rear_nest_trt +
               #rear_nest_LD +
               (1 | rear_mom_Ring) + (1 | rear_dad_Ring) +
               (1 | hatch_year) +
               (1 | rear_area / rear_Box),
             data = nest,))
tab_model(mod7, title = "Table 7")

```

***
Fledging success was significantly lower in enlarged broods; however, it did 
not differ significantly between controls and reduced broods (Table 8).
```{r, warning = F, echo = F, cache = T }
mod8 <- (glmer(
  cbind(number_chicks_fledged_from_rear_nest, died) ~ rear_nest_trt +
    #rear_nest_LD +
    (1 | rear_mom_Ring) + (1 | rear_dad_Ring) +
    (1 | hatch_year) +
    (1 | rear_area / rear_Box),
  data = nest,
  family = "binomial"
))
tab_model(mod8, title = "Table 8")
```

***

On average, enlarged broods failed to fledge ~ 2 nestlings, while mortality 
rates for control and reduced broods was < 1. 
```{r}
aggregate(died ~ rear_nest_trt, data = nest, mean) 
```


### Discussion
Enlarging broods led to worse outcomes both for individual nestlings and for reproductive success of parents. 
Nestlings in enlarged broods tended to be smaller in mass and skeletal size (tarsus length). Mass and tarsus length 
were both significantly negatively associated with brood size. Parents of enlarged broods also
fledged a smaller proportion of their broods compared to controls. On the other hand, nestlings in reduced 
broods were not larger compared to controls, although they did seem more likely to survive to the next year, 
suggesting that they did receive more or better care from their parents. Parents of enlarged broods 
seemed unable or unwilling to increase care, which likely resulted in poorer quality nestlings and lower 
fledging success. 

The variance in quality among nestlings in the enlarged brood was likewise larger compared to controls, consistent 
with the hypothesis that parents of enlarged broods struggled to care for all nestlings equally. It is difficult to determine, however, whether these results demonstrate that nestlings are competing within the brood. Although having more siblings tended to reduce a nestling's condition and likelihood of fledging, that does not necessarily
mean that nestlings were directly competing with one another. Parents may have limited their investment 
in enlarged broods according to their own condition and/or opportunuties for future reproduction, leading 
to poorer quality broods, independent of interactions between individual nestmates. However, the fact that 
variance in quality was higher in reduced broods is consistent with the hypothesis that some nestlings in 
enlarged broods were still managing to grow and develop at the expense of their nestmates. 

Provisioning and begging data could help identify whether parents are feeding nestlings according to need or according to which nestling signals most vigorously. If parents feed the biggest and most active nestlings disproportionately, 
that would be a strong indication that dominant nestlings can outcompete siblings for food. 

### Citation and Session Info
```{r}

citation("lme4")
citation("lmerTest")
sessionInfo()
```
