---
title: "predict_issue"
author: "Sabrina McNew"
date: "12/11/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#### Project info: https://osf.io/34fzc/
#### Contact: sm983@cornell.edu

```{r, echo = F, message=F, cache = TRUE}
library(lme4)
library(lmerTest)
library(dplyr)
library(here)
```

```{r, echo = T, message=F, cache = TRUE}
# load data 
tit <- read.csv(here("data/blue_tit_data_updated_2020-04-18.csv"), stringsAsFactors = FALSE)

# Turn these integer columns into factors
tit <- tit %>%  mutate_at(
  vars(
    hatch_year,
    Extra.pair_paternity,
    rear_nest_trt,
    home_or_away,
    chick_sex_molec,
    chick_survival_to_first_breed_season,
    rear_mom_Ring,
    rear_dad_Ring,
    hatch_mom_Ring,
    genetic_dad_ring_.WP_or_EP.,
    rear_area,
    rear_Box
  ),
  as.factor
)
tit$rear_nest_trt <- relevel(tit$rear_nest_trt, ref = "7")

# Predictor data
# Small but annoying issue with predictor data: warning "incomplete final line found by readTableHeader"
# Solution: save as tsv (.txt), open in text editor, place cursor on last character
# and press "enter."
prediction_data <-read.table(here("data/blue_tit_percentiles_for_supplement_wide.txt"),
                             header =T,
                             stringsAsFactors = FALSE)
prediction_data <- prediction_data  %>%  mutate_at(
                                              vars(
                                                hatch_year,
                                                Extra.pair_paternity,
                                                rear_nest_trt,
                                                home_or_away,
                                                chick_sex_molec,
                                                chick_survival_to_first_breed_season,
                                                rear_mom_Ring,
                                                rear_dad_Ring,
                                                hatch_mom_Ring,
                                                genetic_dad_ring_.WP_or_EP.,
                                                rear_area,
                                                rear_Box
                                                ),
                                              as.factor)
prediction_data$rear_nest_trt <- relevel(prediction_data$rear_nest_trt, ref = "7")

prediction_data_orig <- prediction_data[1:3,] %>% droplevels # original predict data 
prediction_data_fix <- prediction_data[2:4,] %>% droplevels # "corrected" predict data 

```

To illustrate the problem/solution I have loaded in predictor data with 4 rows:
the first three are the ones loaded on OSF, and the fourth is modified by changing
the first entry's hatch box and rear box from SW63 to SW116

```{r, cache = TRUE}
prediction_data_orig[,1:5]
prediction_data_fix[,1:5]

```

Fit the model 
```{r, cache = TRUE}
main_mod <-  lmer(
  day_14_weight ~
    rear_Cs_at_start_of_rearing +
    hatch_nest_LD +
    home_or_away +
    chick_sex_molec +
    (1 | hatch_year) + (1 | rear_nest_breed_ID) +
    (1 | rear_mom_Ring) + (1 | rear_dad_Ring) +
    (1 | hatch_mom_Ring) + (1 | genetic_dad_ring_.WP_or_EP.) +
    (1 | rear_area / rear_Box),
  data = tit 
)
```

Predict the model

```{r, error=TRUE, message = FALSE, cache = TRUE}
predict(main_mod,
        type = "link",
          newdata = prediction_data_orig) # generates error
```

Figure out where the error comes from 
```{r, error = TRUE, message = FALSE, cache = TRUE}
modeldf <- main_mod@frame
list_predictors <- names(modeldf)
# train and test variables are the same class 
sapply(list_predictors, function(x) class(prediction_data_orig[,x]) == class(modeldf[,x])) 
# check whether the test factor levels match training factor levels
sapply(list_predictors, function(x) levels(prediction_data_orig[,x]) %in% levels(modeldf[,x]))

# Looks like there's a problem with the rear_Box levels: 
levels(prediction_data_orig[,"rear_Box"])

## SW63 is not in the training dataset
levels(prediction_data_orig[,"rear_Box"])[3] %in% levels(modeldf[,"rear_Box"])

```

So it looks like SW63 isn't in the training set, (in the model levels). 
I checked on it and it is in the dataset but gets dropped when the model is 
fit because it has some missing data. 

What happens if I swap SW63 with another box that is in the training set? 
```{r, error = TRUE, message = FALSE, cache = TRUE}
predict(main_mod,
        type = "link",
          newdata = prediction_data_fix)
```

How do these differ from telling the predict to accept new levels?
(Similar predictions, compare 1 v. 4)
```{r, error = TRUE, message = FALSE, cache = TRUE}
predict(main_mod,
        type = "link",
          newdata = prediction_data_orig, allow.new.levels =T)
```
